Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2012-04-23 extended_catalog/paper/extended_search.tex
updated text

\newcommand{\degree}{^\circ\xspace}

@@

For sources that are comparable in size to the PSF,
the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  
\newtext{In section XXX, we perform a dedicated monte carlo simulation
to asses possible bias due to incorrectly modeling the spatial structure
of an extended source.}
Therefore, in our
search for extended sources we use only a radially-symmetric uniform
disk spatial model. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.

@@

\newtext{
These previous simulations were preformed assuming an isotropic backgorund.
We performed a second test only

The flux of the source was scaled as
\begin{equation}
  F(\vec{x}) = F(GC) \times \left(\frac{B(\vec{x})}{B(GC)
\end{equation}

.
The benifitted from probing a large range of fluxes and spectral indices
and by testing two energy ranges.
But were unable to assess any possible systematics that may be due to
structured diffuse emission.

Assessing possible difficulties due to having
an incorrect

We performed an additional Monte Carlo test with sources

Figure~\ref{tsext_plane_plot}

We note that this study does not asses issues due to incorrectly modeling the 
galactic diffuse background which are much more difficult to assess. But
this test does show the power of our tools when the model is good (or something).

@@

W44 was simulated with a power-law spectral model. The flux and spectral index of this source
were taken from our re-analysis in this paper which is described in Section~\ref{validate_known}
and summarized in Table~\label{known_extended_sources}. In particualr W44 was simulated
with a flux of $7.12\times10^{-8}$ \phflux in the energy range from 1 \gev to 100 \gev with a spectral
index of 2.66.

@@

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the souce was simulated on
top of the standard galactic and isotropic diffuse background model used
in 2FGL.  This simulation was in the 1 \gev to 100 \gev energy range and
the sources were simulated over the two-year time range included in the
2FGL catalog. We performed 985 identical simulations with different seeds to
see the statistical spread in the fit values.

@@

\clearpage
\begin{deluxetable}{rrrrrr}
\tabletypesize{\normalsize}
\tablecaption{Monte Carlo Spectral Parameters
\label{ts_ext_num_sims}
}
\tablewidth{0pt}
\tablehead{
\colhead{Spectral Index}&
\colhead{Flux\tablenotemark{(a)}}&
\colhead{$N_{1-100\gev}$}&
\colhead{$\langle\ts\rangle_{1-100\gev}$}&
\colhead{$N_{10-100\gev}$}&
\colhead{$\langle\ts\rangle_{10-100\gev}$}\\
\colhead{}&
\colhead{(\phflux)}&
\colhead{}&
\colhead{}&
\colhead{}&
\colhead{}
}

\startdata
      1.5 &  $3\times 10^{-7}$ &           18938 &  22233    &     18938   &    8084     \\
          &          $10^{-7}$ &           19079 &   5827    &     19079   &    2258     \\
          &  $3\times 10^{-8}$ &           19303 &   1276    &     19303   &     541     \\
          &          $10^{-8}$ &           19385 &    303    &     19381   &     142     \\
          &  $3\times 10^{-9}$ &           18694 &     62    &     12442   &      43     \\
      \hline
        2 &          $10^{-6}$ &           18760 &  22101    &     18760   &    3033     \\
          &  $3\times 10^{-7}$ &           18775 &   4913    &     18775   &     730     \\
          &          $10^{-7}$ &           18804 &   1170    &     18803   &     192     \\
          &  $3\times 10^{-8}$ &           18836 &    224    &     15256   &      50     \\
          &          $10^{-8}$ &           17060 &     50    &   \nodata   & \nodata     \\
      \hline
      2.5 &  $3\times 10^{-6}$ &           18597 &  19036    &     18597   &     786    \\
          &          $10^{-6}$ &           18609 &   4738    &     18608   &     208    \\
          &  $3\times 10^{-7}$ &           18613 &    954    &     15958   &      53    \\
          &          $10^{-7}$ &           18658 &    203    &   \nodata   & \nodata    \\
          &  $3\times 10^{-8}$ &           14072 &     41    &   \nodata   & \nodata    \\
      \hline                                                
        3 &          $10^{-5}$ &           18354 &  19466    &     18354   &     215    \\
          &  $3\times 10^{-6}$ &           18381 &   4205    &     15973   &      54    \\
          &          $10^{-6}$ &           18449 &    966    &   \nodata   & \nodata    \\
          &  $3\times 10^{-7}$ &           18517 &    174    &   \nodata   & \nodata    \\
          &          $10^{-7}$ &           13714 &     41    &   \nodata   & \nodata    \\
\enddata

\tablenotetext{(a)}{Integral 100 \mev to 100 \gev flux.}

\tablecomments{
    A list of the spectral models of the simulated point-like sources
    which were tested for extension.  For each model, the number of
    statistically independent simulations and the average value of \ts is
    also tabulated.  
}
\end{deluxetable}

>>>>

\newcommand{\degree}{\ensuremath{^\circ}\xspace}

@@

For sources that are comparable in size to the PSF,
the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  
\newtext{In section \ref{bias_wrong_spatial_model}, we perform a dedicated Monte Carlo simulation
to asses possible bias due to incorrectly modeling the spatial structure
of an extended source.}
Therefore, in our
search for extended sources we use only a radially-symmetric uniform
disk spatial model. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.

@@

\newtext{
These previous simulations spaned the range of detectable fluxes 
over two different energy ranges, but only assuming an isotropic backgorund.
We performed a second simulation to assess any effects that could
be introduces by performing this test in the presence of a highly
structured background.
Our simulation setup was the same as above except that the sources
were simulated on top of and analytized assuming the presence of
the standard Galactic and isotropic diffuse background model used
in 2FGL.

In our simulation, we selected our sources to have a random position on the sky
such that the absolute value of the galactic latitude of the source is
less than 5\degree, probing the most structured areas of the galaxy.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we perform
this test only in the 1 \gev to 100 \gev energy range. 
The source flux was selected so that $\ts\sim50$. This is around 
the significane level for which on might
exepcted to be able to spatially resolve the source. 
We do not test brigher sources since the structured background
should produce the most significant effect for very bright sources.

Because the Galactic diffuse emission is highly structure, the point-source
sensitivity can vary signficantly across the galactic plane.
To account for this, we scaled the flux (for a given spectral index)
by the overal background level according to this formula:
\begin{equation}
  \label{scale_flux_by_background}
  F(\vec{x}) = F(\text{GC}) \times \left(
  \frac{B(\text{GC})}{B(\vec{x})}\right)^{1/2}.
\end{equation}
Here, $\vec{x}$ is the position of the source, $F$ is the integral
flux of the simulated source from 100 \mev to 100 \gev, $F(\text{GC})$
is the same quantity if the source were at the galactic center, $B$
is the integral of the Galactic diffuse and isotropic diffuse emission
model from 1 \gev to 100 \gev and $B(\text{GC})$ is the same quanitty
if the source were at the galactic center.  Table~\ref{ts_ext_num_sims}
tabulates $F(\text{GC})$ and the average \ts for the four spectral models.
For each spectral index, we performed $\sim90,000$ simulations.

Figure~\ref{tsext_plane_plot} shows the cumulative density
of \tsext compared to the $\chi^2_1/2$ distribution of
Equation~\ref{ts_ext_distribution}. For small values of \tsext,
there is reasonable agreement between the Monte Carlo simulation and
theory.  At the highest obtained values of \tsext, there is possibly
some discrepancy.  But in this range there are only limited statistics
because the curve is determined by only a few points.

@@

W44 was simulated with a power-law spectral model. The flux and spectral index of this source
were taken from our re-analysis in this paper which is described in Section~\ref{validate_known}
and summarized in Table~\ref{known_extended_sources}. In particualr W44 was simulated
with a flux of $7.12\times10^{-8}$ \phflux in the energy range from 1 \gev to 100 \gev with a spectral
index of 2.66.

@@

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the souce was simulated on
top of and analyzed assuming the presence of the standard galactic and
isotropic diffuse background model used in 2FGL.  This simulation was
in the 1 \gev to 100 \gev energy range and
the sources were simulated over the two-year time range included in the
2FGL catalog. We performed 985 identical simulations with different seeds to
see the statistical spread in the fit values.

@@

\clearpage
\begin{deluxetable}{rrrrrr}
\tabletypesize{\footnotesize}
\tablecaption{Monte Carlo Spectral Parameters
\label{ts_ext_num_sims}
}
\tablecolumns{6}
\tablewidth{0pt}
\tablehead{
\colhead{Spectral Index}&
\colhead{Flux\tablenotemark{(a)}}&
\colhead{$N_{1-100\gev}$}&
\colhead{$\langle\ts\rangle_{1-100\gev}$}&
\colhead{$N_{10-100\gev}$}&
\colhead{$\langle\ts\rangle_{10-100\gev}$}\\
\colhead{}&
\colhead{(\phflux)}&
\colhead{}&
\colhead{}&
\colhead{}&
\colhead{}
}

\startdata
\multicolumn{6}{c}{Isotropic Background} \\
\hline
      1.5 &  $3\times 10^{-7}$ &           18938 &  22233    &     18938   &    8084     \\
          &          $10^{-7}$ &           19079 &   5827    &     19079   &    2258     \\
          &  $3\times 10^{-8}$ &           19303 &   1276    &     19303   &     541     \\
          &          $10^{-8}$ &           19385 &    303    &     19381   &     142     \\
          &  $3\times 10^{-9}$ &           18694 &     62    &     12442   &      43     \\
      \hline
        2 &          $10^{-6}$ &           18760 &  22101    &     18760   &    3033     \\
          &  $3\times 10^{-7}$ &           18775 &   4913    &     18775   &     730     \\
          &          $10^{-7}$ &           18804 &   1170    &     18803   &     192     \\
          &  $3\times 10^{-8}$ &           18836 &    224    &     15256   &      50     \\
          &          $10^{-8}$ &           17060 &     50    &   \nodata   & \nodata     \\
      \hline
      2.5 &  $3\times 10^{-6}$ &           18597 &  19036    &     18597   &     786    \\
          &          $10^{-6}$ &           18609 &   4738    &     18608   &     208    \\
          &  $3\times 10^{-7}$ &           18613 &    954    &     15958   &      53    \\
          &          $10^{-7}$ &           18658 &    203    &   \nodata   & \nodata    \\
          &  $3\times 10^{-8}$ &           14072 &     41    &   \nodata   & \nodata    \\
      \hline                                                
        3 &          $10^{-5}$ &           18354 &  19466    &     18354   &     215    \\
          &  $3\times 10^{-6}$ &           18381 &   4205    &     15973   &      54    \\
          &          $10^{-6}$ &           18449 &    966    &   \nodata   & \nodata    \\
          &  $3\times 10^{-7}$ &           18517 &    174    &   \nodata   & \nodata    \\
          &          $10^{-7}$ &           13714 &     41    &   \nodata   & \nodata    \\
\cutinhead{Galactic and Isotropic Diffuse Background\tablenotemark{(b)}}
      1.5 &  $3\times 10^{-7}$ &           90741 &     63    &   \nodata   & \nodata     \\
        2 &          $10^{-6}$ &           92161 &     60    &   \nodata   & \nodata     \\
      2.5 &  $3\times 10^{-6}$ &           86226 &     47    &   \nodata   & \nodata    \\
        3 &          $10^{-5}$ &           94412 &     61    &   \nodata   & \nodata    \\
\enddata

\tablenotetext{(a)}{Integral 100 \mev to 100 \gev flux.}
\tablenotetext{(b)}{
\newtext{For the Galactic simulations, the quoted fluxes are
if it was placed in the galactic center and is scaled by Equation~\ref{scale_flux_by_background}.}
}

\tablecomments{
    A list of the spectral models of the simulated point-like sources
    which were tested for extension.  For each model, the number of
    statistically independent simulations and the average value of \ts is
    also tabulated.  
    \newtext{The top rows are the sources simulated on top of an isotropic background and
    the bottom rows are the simulations on top of the Galactic and isotropic
    background.}
}
\end{deluxetable}
